# Week 1: DIYS 1

## Aim:

To refresh your R skills by performing some basic analyses (i.e., descriptive, exploratory, and hypothesis testing ones).

## First Part: Descriptive Analysis

1. Download the files `f.txt` and `m.txt`. They contain information on the number of steps in a day and the body mass index (BMI) for female and male individuals respectively. Open them and explore the first 5 observations for each file.

<span style="color:red">Adjust using the links from GitHub []()</span>

```{r}
# Your code goes here

```
<span style="color:red">For the exercise before publishing the solution</span>


```{r}
# open data
female <- read.table("~/Documents/0_IPZ/2023_2/Leemann-QuantMethods/QuantitativeMethods/QuantitativeMethods/Data/f.txt", header = TRUE, sep = "\t")

# explore data
head(female, 3)

# open data
male <- read.table("~/Documents/0_IPZ/2023_2/Leemann-QuantMethods/QuantitativeMethods/QuantitativeMethods/Data/m.txt", header = TRUE, sep = "\t")

# explore data
head(male, 3)
```

Some key functions in dplyr can be categorized as dealing with columns (e.g., `select`, `mutate`), rows (e.g., `filter`, `distinct`, `arrange`), or groups  (e.g., `group_by`, `summarise`, and `count`). Let's use some of them!

2. Select only the columns 'steps' and 'bmi'. Do it only for the first three observations of the data on females.

```{r}
# It's necessary to restate it in each r code section so the book can be rendered.
library(dplyr)
head(female, 3) %>%
  select(steps, bmi)
```  

3. Select all columns except 'ID'. Do not use `steps` nor `bmi`. Do it only for the first three observations of the data on females. Is the resulting table the same as the previous point? If not, check your answer.

```{r}
library(dplyr)
head(female, 3) %>%
  select(-ID)
```  

Note: to check the documentation of `select`, use `?select` on the console.

4. Use `mutate` to create a new column in the dataframe `female` called `StepsTimesBmi` formed as the product of `steps` and  `bmi`. Show the first three observations for the new variable.

```{r}
library(dplyr)
female <-female %>%
  mutate(StepsTimesBmi= steps * bmi)
  
head(female$StepsTimesBmi,3)
```  

5. Get rid of the column `StepsTimesBmi`. Use `subset`.
 
```{r}
female <- subset(female, select= - StepsTimesBmi)
```     

6. Use filter to find the share of female individuals with a `bmi` higher than 20 and lower than 21.
    
```{r}
library(dplyr)
f20_21 <- female %>%
  filter(bmi>20, bmi<21)

cat("The share of female individuals with a `bmi` higher than 20 and lower than 21 is:", nrow(f20_21)*100/nrow(female), "%\n")
```   

7. Use filter to find the share of female individuals with a `bmi` higher than 20 and lower than 21 while at the same time having less than 14000 `steps`.
    
```{r}
library(dplyr)
fBMI20_21_Step14000 <- female %>%
  filter(bmi>20, bmi<21, steps<14000)

cat("The share of female individuals with a `bmi` higher than 20 and lower than 21 while at the same time having less than 14000 is:", nrow(fBMI20_21_Step14000)*100/nrow(female), "%\n")
``` 

8. Use filter to find the share of male individuals with `ID` number lower than 5 **and** higher than 860. Notice that you can use either `&` between conditions or simply a comma. Could any data set generate a different answer? Why?
    
```{r}
library(dplyr)
m_5_860 <- male %>%
  filter(ID<5 & ID>860)

cat("The share of male individuals with `ID` number lower than 5 AND higher than 860 is:", nrow(m_5_860)*100/nrow(male), "%\n")
```   


9. Use filter to find the share of male individuals with `ID` number lower than 5 **or** higher than 860. Use `|` between conditions. Could any data set generate a different answer? Why?
    
```{r}
library(dplyr)
m_5_or_860 <- male %>%
  filter(ID<5 | ID>860)

cat("The share of male individuals with `ID` number lower than 5 OR higher than 860 is:", nrow(m_5_or_860)*100/nrow(male), "%\n")
```  


10. Use `distinct` to identify the share of male IDs that are unique.
    
```{r}
unique_m_IDs <- male %>%
  distinct(ID)
  
cat("The share of male IDs that are unique is:", nrow(unique_m_IDs)*100/nrow(male), "%\n")
```  

11. Use `arrange` to find the three highest and lowest BMI values for males. Use `slice_head`.
    
```{r}
# Max
top_3_m <- male %>%
  arrange(desc(bmi)) %>%
  slice_head(n = 3)
print(top_3_m)

# Min  
bottom_3_m <- male %>%
  arrange(bmi) %>%
  slice_head(n = 3)
print(bottom_3_m)
```  

    9. group_by summarise count
    
```{r}

```  

2. Are there repeated ids within each data set?

```{r}
# get package
# install.packages("dplyr")
library(dplyr)


# Check for repeated IDs in the female data set. How many are there?
repeated_ids_female <- female %>%
  group_by(ID) %>%
  filter(n() > 1)

cat("Number of repeated IDs in the female data set:", nrow(repeated_ids_female), "\n")


# Check for repeated IDs in the male data set. How many are there?
repeated_ids_male <- male %>%
  group_by(ID) %>%
  filter(n() > 1)

cat("Number of repeated IDs in the male data set:", nrow(repeated_ids_male), "\n")

```




## Second Part: Exploratory VS. Hypothesis-Testing Analysis

Please read the whole instruction before solving the exercise.

Each student will be randomly allocated to either doing the task 1 or 2 (a list containing those numbers will published). Both tasks are based on the same data sets used in the first part.

Notes: 

- The details of the data origin will be published with the solution.
- Students allocated to each group are encouraged to do the task for the other group **_only_** after finishing their own task.

### Preliminary steps: do this before doing the task that you were assigned to

1. For each data set, create a new variable called `sex`. Assign any value to each case, but make sure they are different.

```{r}
female$sex <- 'F'  
male$sex <- 'M'  
```
  
2. Create one data frame with all the IDs present in **both** data sets. How many cases are there? Use `dplyr`'s [join](https://www.geeksforgeeks.org/joining-data-in-r-with-dplyr-package/) methods.



```{r}
library(dplyr)

in_both <- inner_join(female, male, by="ID")

cat("The number of cases where an ID is in both data sets is:", nrow(in_both), "\n")

```


3. Now that you know that there are no repeated individuals across the data sets, consider whether a join method is the appropriate way of unifying both data sets. Try first with `full_join` and then with `bind_rows`. Which one should you use? Why? Finally, how many individuals does the new dataframe have?

```{r}

library(dplyr)

all <- full_join(female, male, by="ID", copy=FALSE)
cat("The new dataframe has ", nrow(all), "individuals\n")

# Assuming that `sex` was created for each dataframe
all <- bind_rows(female, male, .id = NULL)
cat("The new dataframe has ", nrow(all), "individuals\n")

# Without assuming that `sex` was created for each dataframe
female <- read.table("~/Documents/0_IPZ/2023_2/Leemann-QuantMethods/QuantitativeMethods/QuantitativeMethods/Data/f.txt", header = TRUE, sep = "\t")
male <- read.table("~/Documents/0_IPZ/2023_2/Leemann-QuantMethods/QuantitativeMethods/QuantitativeMethods/Data/m.txt", header = TRUE, sep = "\t")

all <- bind_rows(female, male, .id = 'sex')
cat("The new dataframe has ", nrow(all), "individuals\n")

# Which assigns a number 1 for the first binded dataframe, and 2 for the second one. Hence, we can replace the values as follows.
all$sex <- ifelse(all$sex == 1, 'F', ifelse(all$sex == 2, 'M', all$sex))

```


4. What's the share per sex in the unified dataframe from the previous point? Use ``





Consider using the packages `dplyr`, ``



### Task 1:
 - What do you conclude from the combined data set (i.e., the one formed using both the one for males and the one for females)?
  - What questions did you ask yourself?
    - Why did you ask those questions? Is there an intuition behind them?
      - If so, what was your intuition?
      - If not, how did you proceed?

- Hint: consider visualizing how variables interact.


### Task 2:

- Is the average number of steps for males and females statistically different?
- How do BMI and daily steps statistically relate to each other?
  - Does that relationship depend on whether individuals are of one sex or another? If so, how?
    - Is there an statistically significant negative correlation between the number of steps and the BMI for females?
    - Is there an statistically significant positive correlation between the number of steps and the BMI for males?
  






- 1st weeks, dplier: to check> to statistical analysis
	- Doing basic code to make analysis (which is fine enough), but in dplier you could do it like this.
	- Make descriptive statistics using an interesting

looking for something unknown in the dark, grope, feel blindly and make conjectures on what things are and how they are related.
- Two groups: random selection: description similar? The smaller the group, the likelier that a random selection is not balanced? What about attrition?

Looking!=seeing:
  Different beliefs (non- and knowledge ones), different preferences, different attention focus -> different attention investment and emphasis
    Value of diverse academic community while keeping a minimal set of shared assessment rules: [objectivity](https://plato.stanford.edu/entries/scientific-objectivity/) as continuum of increasing inter-subjective agreement
    



## Graph
```{r}
# Install and load the ggplot2 package if you haven't already
#install.packages("ggplot2")
library(ggplot2)

# Assuming you have a consolidated dataset named 'combined_data'

# Create a scatterplot of steps vs bmi

ggplot(female, aes(x = steps, y = bmi)) +
  geom_point() +
  labs(x = "Steps", y = "BMI") +
  ggtitle("Scatterplot of Steps vs BMI by Sex") 


ggplot(male, aes(x = steps, y = bmi)) +
  geom_point() +
  labs(x = "Steps", y = "BMI") +
  ggtitle("Scatterplot of Steps vs BMI by Sex") 


  
ggplot(all, aes(x = steps, y = bmi, color = sex)) +
  geom_point() +
  labs(x = "Steps", y = "BMI") +
  ggtitle("Scatterplot of Steps vs BMI by Sex") +
  scale_color_manual(values = c("F" = "blue", "M" = "red"))  # Optional: Define color mapping

```




## Solution

Will be made available.

<!-- Here is the solution -->




